

%---------------------------------------------------------------------------------------------%
<p>
####        * {Dimensions of a data set}
We have remarked that some data sets are very large. This is perhaps a good place to consider summary information about data objects.
For a simple vector – a useful command to determine the length (remark: sample size) is the function texttt{length()}.
<pre>
<code>
> Y=5:9
> length(Y)
[1] 5
</code>
</pre>
<p>
%---------------------------------------------------------------------------------------------%
For more complex data sets (and data frames, which we will see later) – we have several tools for assessing the size of data.
<pre>
<code>
> dim(iris)  # dimensions of data set
[1] 150   5
> nrow(iris) # number of rows
[1] 150
> ncol(iris) # number of columns
[1] 5
</code>
</pre>
<p>
%---------------------------------------------------------------------------------------------%
We can also determine the row names and column names using the functions texttt{rownames()} and texttt{colnames()}.
If there are no specific row or column names – the command will just return the indices.

<pre>
<code>
> colnames(iris)
[1] "Sepal.Length" "Sepal.Width"  "Petal.Length" "Petal.Width"  "Species"

</code>
</pre>
<p>
The command texttt{summary()} is one of the most useful commands in R. It is a generic function used to produce result summaries of the results of various functions. The function invokes particular methods which depend on the class of the first argument. 
In other words – texttt{R} picks out the most suitable type of summary for that data.
{
	small
<pre>
<code>
> summary(iris)
  Sepal.Length    Sepal.Width     Petal.Length    Petal.Width          Species  
 Min.   :4.300   Min.   :2.000   Min.   :1.000   Min.   :0.100   setosa    :50  
 1st Qu.:5.100   1st Qu.:2.800   1st Qu.:1.600   1st Qu.:0.300   versicolor:50  
 Median :5.800   Median :3.000   Median :4.350   Median :1.300   virginica :50  
 Mean   :5.843   Mean   :3.057   Mean   :3.758   Mean   :1.199                  
 3rd Qu.:6.400   3rd Qu.:3.300   3rd Qu.:5.100   3rd Qu.:1.800                  
 Max.   :7.900   Max.   :4.400   Max.   :6.900   Max.   :2.500                  
>
</code>
</pre>
<p>
}

