---
title: "``ModelMetrics::plot_counts``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ModelMetrics)
```

Package ‘ModelMetrics’
November 3, 2018
Title Rapid Calculation of Model Metrics
Version 1.2.2
Date 2018-11-03

#### Description  
 Collection of metrics for evaluating models written in C++ using 'Rcpp'. Popular metrics
include area under the curve, log loss, root mean square error, etc.
Depends R (>= 3.2.2)
License GPL (>= 2)
Encoding UTF-8
LazyData true
LinkingTo Rcpp
Imports Rcpp, data.table
RoxygenNote 6.0.1
Suggests testthat
NeedsCompilation yes
Author Tyler Hunt [aut, cre]
Maintainer Tyler Hunt <thunt@snapfinance.com>
  Repository CRAN
Date/Publication 2018-11-03 17:00:14 UTC
R topics documented:
  auc . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 2
brier . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 3
ce . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 4
confusionMatrix . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 5
f1Score . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 5
fScore . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 6
gini . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 6
kappa . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 7
logLoss . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 8
1
2 auc
mae . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 9
mauc . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 9
mcc . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 10
mlogLoss . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 11
mse . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 11
msle . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 12
npv . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 13
ppv . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 13
recall . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 14
rmse . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 15
rmsle . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 15
testDF . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 16
tnr . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 17
Index 18
auc Area Under the Curve

#### Description  

Calculates the area under the curve for a binary classifcation model

#### Usage  
<pre><code>

auc(...)
## Default S3 method:
auc(actual, predicted, ...)
## S3 method for class 'glm'
auc(modelObject, ...)
## S3 method for class 'randomForest'
auc(modelObject, ...)
## S3 method for class 'glmerMod'
auc(modelObject, ...)
## S3 method for class 'gbm'
auc(modelObject, ...)
## S3 method for class 'rpart'
auc(modelObject, ...)
brier 3

</code></pre>
#### Arguments

... additional parameters to be passed the the s3 methods
* ``actual``: A vector of the labels. Can be numeric, character, or factor
* ``predicted``:  A vector of predicted values
* ``modelObject``:  the model object. Currently supported glm, randomForest, glmerMod, gbm

#### Examples  

data(testDF)
glmModel <- glm(y ~ ., data = testDF, family="binomial")
Preds <- predict(glmModel, type = 'response')
auc(testDF$y, Preds)
# using s3 method for glm
auc(glmModel)
brier Brier Score

#### Description  

Calculates the Brier score

#### Usage  
<pre><code>

brier(...)
## Default S3 method:
brier(actual, predicted, ...)
## S3 method for class 'glm'
brier(modelObject, ...)
## S3 method for class 'randomForest'
brier(modelObject, ...)
## S3 method for class 'glmerMod'
brier(modelObject, ...)
## S3 method for class 'gbm'
brier(modelObject, ...)
## S3 method for class 'rpart'
brier(modelObject, ...)
4 ce

</code></pre>
#### Arguments

... additional parameters to be passed the the s3 methods
* ``actual``: A vector of the labels
* ``predicted``:  A vector of predicted values
* ``modelObject``:  the model object. Currently supported glm, randomForest, glmerMod, gbm
ce Classification error

#### Description  

Calculates the classification error

#### Usage  
<pre><code>

ce(...)
## Default S3 method:
ce(actual, predicted, ...)
## S3 method for class 'lm'
ce(modelObject, ...)
## S3 method for class 'glm'
ce(modelObject, ...)
## S3 method for class 'randomForest'
ce(modelObject, ...)
## S3 method for class 'glmerMod'
ce(modelObject, ...)
## S3 method for class 'gbm'
ce(modelObject, ...)
## S3 method for class 'rpart'
ce(modelObject, ...)

</code></pre>
#### Arguments

... additional parameters to be passed the the s3 methods
* ``actual``: A vector of the labels
* ``predicted``:  A vector of predicted values
* ``modelObject``:  the model object. Currently supported lm, glm, randomForest, glmerMod, gbm, rpart
confusionMatrix 5
confusionMatrix Confusion Matrix

#### Description  

Create a confusion matrix given a specific cutoff.

#### Usage  
<pre><code>

confusionMatrix(actual, predicted, cutoff = 0.5)

</code></pre>
#### Arguments

* ``actual``: A vector of the labels
* ``predicted``:  A vector of predicted values
cutoff A cutoff for the predicted values
f1Score F1 Score

#### Description  

Calculates the f1 score

#### Usage  
<pre><code>

f1Score(actual, predicted, cutoff = 0.5)

</code></pre>
#### Arguments

* ``actual``: A vector of the labels
* ``predicted``:  A vector of predicted values
cutoff A cutoff for the predicted values

fScore F Score

#### Description  

Calculates the F score and allows different specifications of the beta value (F0.5)

#### Usage  
<pre><code>

fScore(actual, predicted, cutoff = 0.5, beta = 1)

</code></pre>
#### Arguments

* ``actual``: A vector of the labels
* ``predicted``:  A vector of predicted values
cutoff A cutoff for the predicted values
beta the desired beta value (lower increases weight of precision over recall). Defaults
to 1
gini GINI Coefficient

#### Description  

Calculates the GINI coefficient for a binary classifcation model

#### Usage  
<pre><code>

gini(...)
## Default S3 method:
gini(actual, predicted, ...)
## S3 method for class 'glm'
gini(modelObject, ...)
## S3 method for class 'randomForest'
gini(modelObject, ...)
## S3 method for class 'glmerMod'
gini(modelObject, ...)
## S3 method for class 'gbm'
gini(modelObject, ...)
## S3 method for class 'rpart'
gini(modelObject, ...)
kappa 7

</code></pre>
#### Arguments

... additional parameters to be passed the the s3 methods
* ``actual``: A vector of the labels. Can be numeric, character, or factor
* ``predicted``:  A vector of predicted values
* ``modelObject``:  the model object. Currently supported glm, randomForest, glmerMod, gbm

#### Examples  

data(testDF)
glmModel <- glm(y ~ ., data = testDF, family="binomial")
Preds <- predict(glmModel, type = 'response')
gini(testDF$y, Preds)
# using s3 method for glm
gini(glmModel)
kappa kappa statistic

#### Description  

Calculates kappa statistic. Currently build to handle binary values in actual vector.

#### Usage  
<pre><code>

kappa(actual, predicted, cutoff = 0.5)

</code></pre>
#### Arguments

* ``actual``: A vector of the labels
* ``predicted``:  A vector of predicted values
cutoff A cutoff for the predicted values
8 logLoss
logLoss Log Loss

#### Description  

Calculates the log loss or entropy loss for a binary outcome

#### Usage  
<pre><code>

logLoss(...)
## Default S3 method:
logLoss(actual, predicted, distribution = "binomial", ...)
## S3 method for class 'glm'
logLoss(modelObject, ...)
## S3 method for class 'randomForest'
logLoss(modelObject, ...)
## S3 method for class 'glmerMod'
logLoss(modelObject, ...)
## S3 method for class 'gbm'
logLoss(modelObject, ...)
## S3 method for class 'rpart'
logLoss(modelObject, ...)

</code></pre>
#### Arguments

... additional parameters to be passed the the s3 methods
actual a binary vector of the labels
predicted a vector of predicted values
distribution the distribution of the loss function needed binomial, poisson
* ``modelObject``:  the model object. Currently supported glm, randomForest, glmerMod, gbm

#### Examples  

data(testDF)
glmModel <- glm(y ~ ., data = testDF, family="binomial")
Preds <- predict(glmModel, type = 'response')
logLoss(testDF$y, Preds)
# using s3 method for glm
logLoss(glmModel)
mae 9
mae Mean absolute error

#### Description  

Calculates the mean absolute error

#### Usage  
<pre><code>

mae(...)
## Default S3 method:
mae(actual, predicted, ...)
## S3 method for class 'glm'
mae(modelObject, ...)
## S3 method for class 'randomForest'
mae(modelObject, ...)
## S3 method for class 'glmerMod'
mae(modelObject, ...)
## S3 method for class 'gbm'
mae(modelObject, ...)
## S3 method for class 'rpart'
mae(modelObject, ...)

</code></pre>
#### Arguments

... additional parameters to be passed the the s3 methods
* ``actual``: A vector of the labels
* ``predicted``:  A vector of predicted values
* ``modelObject``:  the model object. Currently supported glm, randomForest, glmerMod, gbm
* ``mauc``: Multiclass Area Under the Curve

#### Description  

Calculates the area under the curve for a binary classifcation model
10 mcc

#### Usage  
<pre><code>

mauc(actual, predicted)

</code></pre>
#### Arguments

* ``actual``: A vector of the labels. Can be numeric, character, or factor
* ``predicted``: A data.frame of predicted values. Can be matrix, data.frame

#### Examples  

```{r}
setosa <- glm(I(Species == 'setosa') ~ Sepal.Length, data = iris, family = 'binomial')
versicolor <- glm(I(Species == 'versicolor') ~ Sepal.Length, data = iris, family = 'binomial')
virginica <- glm(I(Species == 'virginica') ~ Sepal.Length, data = iris, family = 'binomial')
Pred <-
  data.frame(
    setosa = predict(setosa, type = 'response')
    ,versicolor = predict(versicolor, type = 'response')
    ,virginica = predict(virginica, type = 'response')
  )
Predicted = Pred/rowSums(Pred)
Actual = iris$Species
mauc(Actual, Predicted)
```

mcc Matthews Correlation Coefficient

#### Description  

Calculates the Matthews Correlation Coefficient

#### Usage  
<pre><code>

mcc(actual, predicted, cutoff)

</code></pre>
#### Arguments

* ``actual``: A vector of the labels
* ``predicted``:  A vector of predicted values
cutoff A cutoff for the predicted values
mlogLoss 11
mlogLoss Multiclass Log Loss

#### Description  

Calculated the multi-class log loss

#### Usage  
<pre><code>

mlogLoss(actual, predicted)

</code></pre>
#### Arguments

* ``actual``: A vector of the labels. Can be numeric, character, or factor
predicted matrix of predicted values. Can be matrix, data.frame
* ``mse ``:  Mean Square Error

#### Description  

Calculates the mean square error

#### Usage  
<pre><code>

mse(...)
## Default S3 method:
mse(actual, predicted, ...)
## S3 method for class 'lm'
mse(modelObject, ...)
## S3 method for class 'glm'
mse(modelObject, ...)

</code></pre>
#### Arguments

... additional parameters to be passed the the s3 methods
* ``actual``: A vector of the labels
* ``predicted``:  A vector of predicted values
* ``modelObject``:  the model object. Currently supported lm
12 msle

#### Examples  

data(testDF)
glmModel <- glm(y ~ ., data = testDF, family="binomial")
Preds <- predict(glmModel, type = 'response')
mse(testDF$y, Preds)



