mtcars - Regression Exercise
========================================================

### Quick Inspection of Data Set

Firstly we have a quick loot at the data set to set a sense of the 
size of the data set, and the names of the variables
```{r}
tail(mtcars)

```

To find out more about this data set, simpy type `help(mtcars)` to get the help file.

### Exercise 1 : Simple Linear Regression

In this exercise, fit a simple linear model, where weight (***wt***) is the independent variable, and miles per gallon (***mpg***) is the dependent variable. Call the fitted model `Fit1`.
```{r fig.width=7, fig.height=6}

Fit1 = lm(mpg ~ wt,data=mtcars)
Fit1
```
More information about the model can be found using the `summary()` command.

```{r fig.width=7, fig.height=6}

summary(Fit1)

```
### Exercise 2 : Multiple Linear Regression
In this exercise, fit a simple linear model, where weight (***wt***) and number of cylinders (***cyl*** are the independent variable, and miles per gallon (***mpg***) is the dependent variable. Call the fitted model `Fit2`.
```{r fig.width=7, fig.height=6}

Fit2 = lm(mpg ~ wt+cyl,data=mtcars)
Fit2
summary(Fit2)
```

### Exercise 3 : Confidence Intervals for Regression Estimates
Using the regression model from the previous exercise (i.e. `Fit2`), compute 95% and 99% confidence intervals for the regression estimates.

```{r fig.width=7, fig.height=6}
confint(Fit2)
confint(Fit2, conf.level=0.99)
```

### Exercise 4 : Akaike Information Criterion
When comparing two or more candidate models, the model with lowest
AIC value is considered the best model. Use the AIC to compare `Fit1` and `Fit2`.
```{r fig.width=7, fig.height=6}
AIC(Fit1)
AIC(Fit2)
```

### Exercise 5 : ANOVA tables for Regression Models
Construct the ANOVA tables for the fitted models computed previously.
```{r fig.width=7, fig.height=6}
anova(Fit1)
anova(Fit2)
```