	\chapter { R Graphics}
	
	\section{Important Graphical Procedures}
	\begin{enumerate}
		\item Histograms
		\item Box-plots
		\item Scatter-plots
	\end{enumerate}
	
	
	%----------------------------------------------------------------%
	
	
	
			\section Enhancing your scatter plots
			\subsection{Adding lines}
			Previously we have used scatter plots to plot bivariate data. They were constructed using the plot() command.
			Recall that we can use the arguments \texttt{xlim} and \texttt{ylim} to control the vertical and horizontal range of the plots, by specifying a two element vector (min and max) for each.
			
			Using the \texttt{abline()} command, we can add lines to our scatter plots. We specify the argument according to the type of line required. A demonstration of three types of line is provided below.
			Additionally we change the colour of the added lines, by specifying a colour in the \texttt{col} argument. We can also change the line type to one of four possible types, using the \texttt{lty} argument.
			

			
			The \texttt{abline()} function can be used to add a regression model line  by supplying as an argument the \texttt{coef()} values for intercept and slope estimates .These estimates can be inputted directly by using both functions in conjunction.
			
			\footnotesize \begin{verbatim}
			Fit1 =lm(y1~x);  coef(Fit1)
			abline(coef(Fit1))	
			\end{verbatim}\normalsize
			
			\subsection{Adding a title }
			
			It is good practice to label your scatterplots properly. You can specify the following argument
			\begin{itemize}
				\item	main="Scatterplot Example", 	This provides the plot with a title
				\item	sub="Subtitle",                 This adds a subtitle
				\item	xlab="X variable ",				This command labels the x axis 
				\item   ylab="y variable ",				This command labels the y-axis
			\end{itemize}
			We can also add text to each margin, using the \texttt{mtext()} command.  
			We simply require the number of the side. (1 = bottom, 2=left,3=top,4=right). 
			We can change the colour using the col argument.
			\footnotesize \begin{verbatim}
			plot(x,y,main="Scatterplot Example",   sub="subtitle",    xlab="X variable ", ylab="y variable ")	
			mtext("Enhanced Scatterplot", side=4,col="red ")
			\end{verbatim}\normalsize
			Alternatively , we can also use the command title() to add a title to an existing scatterplot.
			\footnotesize \begin{verbatim}
			title(main="Scatterplot Example)	
			\end{verbatim}\normalsize
			
			
			\section{Combining plots}
			It is possible to combine two plots. We used the graphical parameters command \texttt{par()} to create an array. 
			Often we just require two plots side by side or above and below. We simply specify the numbers of rows and columns of this array using the \texttt{mfrow} argument, passed as a vector.
			
			\begin{verbatim}
			par(mfrow=c(1,2))
			plot(x,y1)			# draw first plot
			plot(x,y2)			# draw second plot
			par(mfrow=c(1,1))		# reset to default setting.
			\end{verbatim}
			
			\section{Plot of single vectors}
			If only one vector is specified i.e. \texttt{plot(x)},  the plot created will simply be a scatter-plot of the values of x against their indices.
			
			$plot(x)$
			Suppose we wish to examine a trend that these points represent. We can connect each covariate using a line.
			
			$plot(x, type = "l")$
			If we wish to have both lines and points, we would input the following code. This is quite useful if we wish to see how a trend develops over time.
			$plot(x, type = "b")$
			
			
			
			\section{Exercise}
			
			Generate a histogram for data set 'scores', with an accompanying box-and-whisker plot.
			The colour of the histogram's bar should be yellow. The orientation for the boxplot should be horizontal.
			
			\begin{verbatim}
			scores <-c(23,19,22,22,19,20,25,26,26,19,24,23,17,21,28,26)
			
			par(mfrow=c(2,1)) 	# two rows , one column
			
			hist(scores,main="Distribution of scores",xlab="scores",col="yellow")
			
			boxplot(scores ,horizontal=TRUE)
			
			par(mfrow =c(1,1)) 	#reset
			\end{verbatim}
