{arsenal} {data-navmenu="arsenal"}
========================================

```{r include=FALSE}

library(arsenal)
```



Column {.tabset}
-------------------------------

### ``arsenal::tableby.internal``


Helper functions for tableby

#### Description

A set of helper functions for tableby.

#### Usage
<pre><code>

is.tableby(x)
is.summary.tableby(x)
modpval.tableby(x, pdata, use.pname = FALSE)
tests(x)
## S3 method for class 'tableby'
tests(x)
na.tableby(lhs = TRUE)
## S3 method for class 'tableby'
xtfrm(x)
## S3 method for class 'tableby'
sort(x, ...)
## S3 method for class 'tableby'
Ops(e1, e2)
## S3 method for class 'tableby'
head(x, n = 6L, ...)
## S3 method for class 'tableby'
tail(x, n = 6L, ...)42
tableby.stats

</code></pre>
#### Arguments

* ``x``:  A tableby object.
* ``pdata``: A named data.frame where the first column is the by-variable names, the (optional) second is the strata value, the next is the x variable names, the next is
p-values (or some test stat), and the (optional) next column is the method name.
use.pname Logical, denoting whether the column name in pdata corresponding to the p-values should be used in the output of the object.
lhs Logical, denoting whether to remove NAs from the first column of the data.frame
(the "left-hand side")
... Other arguments, or a vector of indices for extracting.
e1, e2 tableby objects, or numbers to compare them to.
* ``n``:   A single integer. See head or tail for more details
Details
Logical comparisons are implemented for Ops.tableby.

#### Value

na.tableby returns a subsetted version of object (with attributes). Ops.tableby returns a logical
vector. xtfrm.tableby returns the p-values (which are ordered by order to sort).
See Also
arsenal_table, sort, head, tail

<p>

### ``arsenal::summary.tableby``

The summary method for a tableby object

#### Description

The summary method for a tableby object, which is a pretty rendering of a tableby object into a publication-quality results table in R Markdown, and can render well in text-only.

#### Usage
<pre><code>

## S3 method for class 'tableby'
summary(object, ..., labelTranslations = NULL,
text = FALSE, title = NULL, pfootnote = FALSE, term.name = "")
## S3 method for class 'summary.tableby'
as.data.frame(x, ..., text = x$text,
pfootnote = x$pfootnote, term.name = x$term.name, width = NULL,
min.split = NULL, list.ok = FALSE)

</code></pre>
#### Arguments

object : An object of class "tableby", made by the tableby function.
... : For summary.tableby, other arguments passed to as.data.frame.tableby.
For printing the summary object, these are passed to both as.data.frame.summary.tableby
and kable.
labelTranslations : A named list (or vector) where the name is the label in the output to be replaced in the pretty rendering by the character string value for the named element of
the list, e.g., list(age = "Age(Years)", meansd = "Mean(SD)").
text An argument denoting how to print the summary to the screen. Default is FALSE (show markdown output). TRUE and NULL output a text-only version, with the latter avoiding all formatting. "html" uses the HTML tag <strong> instead of the markdown formatting, and "latex" uses the LaTeX command \textbf.
title Title for the table, defaulting to NULL (no title)
pfootnote Logical, denoting whether to put footnotes describing the tests used to generate the p-values.
term.name A character vector denoting the column name for the "terms" column. It should be the same length as the number of tables or less (it will get recycled if needed).
The special value TRUE will use the y-variable’s label for each table.
x An object of class "summary.tableby".
width Passed to smart.split for formatting of the "term" column.
min.split Passed to smart.split for formatting of the "term" column.
list.ok If the object has multiple by-variables, is it okay to return a list of data.frames instead of a single data.frame? If FALSE but there are multiple by-variables, a warning is issued.

#### Value

An object of class summary.tableby
Author(s)
Ethan Heinzen, based on code by Gregory Dougherty, Jason Sinnwell, Beth Atkinson, adapted from
SAS Macros written by Paul Novotny and Ryan Lennon
See Also
tableby.control, tableby

#### Examples
```{r}
set.seed(100)
## make 3+ categories for response
nsubj <- 90
mdat <- data.frame(Response=sample(c(1,2,3),nsubj, replace=TRUE),
Sex=sample(c("Male", "Female"), nsubj,replace=TRUE),
Age=round(rnorm(nsubj,mean=40, sd=5)),
HtIn=round(rnorm(nsubj,mean=65,sd=5)))
## allow default summaries on RHS variables
out <- tableby(Response ~ Sex + Age + HtIn, data=mdat)
summary(out, text=TRUE)
labels(out)
labels(out) <- c(Age="Age (years)", HtIn="Height (inches)")
summary(out, stats.labels=c(meansd="Mean-SD", q1q3 = "Q1-Q3"), text=TRUE)
```
<p>

### "``arsenal::summary.modelsum``"



Summarize a modelsum object.

#### Description

Format the information in object as a table using Pandoc coding or plain text, and cat it to stdout.

#### Usage
<pre><code>

## S3 method for class 'modelsum'
summary(object, ..., labelTranslations = NULL,
text = FALSE, title = NULL, term.name = "")
## S3 method for class 'summary.modelsum'
as.data.frame(x, ..., text = x$text,
term.name = x$term.name, width = NULL, min.split = NULL,
list.ok = FALSE)

</code></pre>
#### Arguments

object
...
A modelsum object.
For summary.modelsum, other arguments passed to as.data.frame.modelsum.
For as.data.frame.summary.modelsum, "width" and "min.split" are passed
to smart.split. For printing the summary object, these are passed to both
as.data.frame.summary.modelsum and kable.
labelTranslations
A named list (or vector) where the name is the label in the output to be replaced
in the pretty rendering by the character string value for the named element of
the list, e.g., list(age = "Age(Years)", meansd = "Mean(SD)").34
summary.tableby

* text An argument denoting how to print the summary to the screen. Default is FALSE
(show markdown output). TRUE and NULL output a text-only version, with the
latter avoiding all formatting. "html" uses the HTML tag <strong> instead of
the markdown formatting, and "latex" uses the LaTeX command \textbf.
* title Title for the table, defaulting to NULL (no title)
* term.name A character vector denoting the column name for the "terms" column. It should be the same length as the number of tables or less (it will get recycled if needed).
The special value TRUE will use the y-variable’s label for each table.
* x An object of class "summary.modelsum".
* width Passed to smart.split for formatting of the "term" column.
* min.split Passed to smart.split for formatting of the "term" column.
* list.ok If the object has multiple by-variables, is it okay to return a list of data.frames instead of a single data.frame? If FALSE but there are multiple by-variables, a
warning is issued.

#### Value

An object of class "summary.modelsum"
Author(s)
Ethan Heinzen, based on code originally by Greg Dougherty
See Also
modelsum, as.data.frame.modelsum
