---
title: "``flashlight::ale_profile``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Index 47
ale_profile ALE profile

#### Description

Internal function used by light_profile to calculate ALE profiles.
ale_profile 3

#### Usage
<pre><code>

ale_profile(x, v, breaks = NULL, n_bins = 11, cut_type = c("equal",
"quantile"), value_name = "value", counts_name = "counts",
counts = TRUE, counts_weighted = FALSE, pred = NULL,
evaluate_at = NULL, indices = NULL, n_max = 1000, seed = NULL,
two_sided = FALSE, calibrate = TRUE)
Arguments
* ``x``: An object of class flashlight.
v The variable to be profiled.
breaks Cut breaks for a numeric v. Only used if no evaluate_at is specified.
* ``n_bins``: Maxmium number of unique values to evaluate for numeric v. Only used if no
evaluate_at is specified.
cut_type For the default "equal", bins of equal width are created for v by pretty. Choose
"quantile" to create quantile bins.
value_name Column name containing the profile value. Defaults to "value".
* ``counts_name``:  Name of the column containing counts if counts is TRUE.
counts Should counts be added?
counts_weighted
If counts is TRUE: Should counts be weighted by the case weights? If TRUE,
the sum of w is returned by group.
pred Optional vector with predictions.
evaluate_at Vector with values of v used to evaluate the profile. Only relevant for type =
"partial dependence".
indices A vector of row numbers to consider.
* ``n_max``:  Maximum number of ICE profiles to calculate within interval (not within data).
seed Integer random seed passed to light_ice.
two_sided Standard ALE profiles are calculated via left derivatives. Set to TRUE if twosided
derivatives should be calculated. Only works for continuous v. More
specifically: Usually, local effects at value x are calculated using points between
x-e and x. Set ale_two_sided = TRUE to use points between x-e/2 and x+e/2.
calibrate Should values be calibrated based on average preditions? Default is TRUE.

#### Value

A tibble containing results.
4 auto_cut
all_identical all_identical

#### Description

Checks if an aspect is identical for all elements in a nested list. The aspect is specified by fun, e.g.
‘[[‘, followed by the element name to compare.

#### Usage
<pre><code>

all_identical(x, fun, ...)
Arguments
x A nested list of objects.
fun Function used to extract information of each element of x.
... Further arguments passed to fun.

#### Value

A logical vector of length one.

#### Examples
```{r}

x <- list(a = 12)
y <- list(a = 13)
all_identical(list(x, y), `[[`, "a")
all_identical(list(x, y), `[[`, "b")
```


