---
title: "``flashlight::plot_counts``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
``auto_cut`` Discretizes a Vector

#### Description

This function takes a vector x and returns a list with information on disretized version of x, see
return for details on the resulting object.

#### Usage
<pre><code>

auto_cut(x, breaks = NULL, n_bins = 27, cut_type = c("equal",
"quantile"), x_name = "value", level_name = "level", ...)
auto_cut 5
</code></pre>

#### Arguments
x A vector.
* ``breaks``: An optional vector of breaks. Only relevant for numeric x.
* ``n_bins``:  If x is numeric and no breaks are provided, this is the maximum number of bins allowed or to be created (approximately).
cut_type For the default type "equal", bins of equal width are created by pretty. Choose "quantile" to create quantile bins.
x_name Column name with the values of x in the output.
level_name Column name with the bin labels of x in the output.
... Further arguments passed to cut3.

#### Details

The construction of level names can be controlled by passing ... arguments to formatC.

#### Value

A list with the following four elements:
*  data A data.frame with colums x_name and level_name each with the same length as x.
The column x_name has values in output bin_means while the column level_name has values in bin_labels.
*  breaks A vector of increasing and unique breaks used to cut a numeric x with too many
distinct levels. NULL otherwise.
*  bin_means The midpoints of subsequent breaks, or if there are no breaks in the output, factor
levels or distinct values of x.
*  bin_labels Break labels of the form "(low, high]" if there are breaks in the output, otherwise
the same as bin_means. Same order as bin_means.

#### Examples
```{r}

auto_cut(1:10, n_bins = 3)
auto_cut(c(NA, 1:10), n_bins = 3)
auto_cut(1:10, breaks = 3:4, n_bins = 3)
auto_cut(1:10, n_bins = 3, cut_type = "quantile")
auto_cut(LETTERS[4:1], n_bins = 2)
auto_cut(factor(LETTERS[1:4], LETTERS[4:1]), n_bins = 2)
auto_cut(990:1100, n_bins = 3, big.mark = "'", format = "fg")
auto_cut(c(0.0001, 0.0002, 0.0003, 0.005), n_bins = 3, format = "fg")
```

