---
title: "``flashlight::plot_counts``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
cut3 Modified cut

#### Description

Slightly modified version of base::cut.default. Both modifications refer to the construction of break labels. Firstly, ... arguments are passed to formatC in formatting the numbers in the labels. Secondly,
a separator between the two numbers can be specified with default ", ".

#### Usage
<pre><code>

cut3(x, breaks, labels = NULL, include.lowest = FALSE, right = TRUE,
dig.lab = 3L, ordered_result = FALSE, sep = ", ", ...)
</code></pre>

#### Arguments
* ``x``: Numeric vector.
breaks Numeric vector of cut points or a single number specifying the number of intervals desired.
labels Labels for the levels of the final categories.
include.lowest Flag if minimum value should be added to intervals of type (,] (or maximum for
[, )).
right Flag if intervals should be closed to the right or left.
dig.lab Number of significant digits passed to formatC.
ordered_result Flag if resulting output vector should be ordered.
sep Separater between from-to labels.
... Arguments passed to formatC.
</code></pre>

#### Value

Vector of the same length as x.

#### Examples
```{r}

x <- 998:1001
cut3(x, breaks = 2)
cut3(x, breaks = 2, big.mark = "'", sep = ":")
```
