---
title: "``flashlight::grouped_stats``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 Grouped Weighted Means or Quartiles

#### Description

Calculates weighted means or quartiles (and counts) of a variable grouped by optional columns.

#### Usage
<pre><code>

grouped_stats(data, xNULL, by = NULL, stats = c("mean",
"quartiles"), counts = TRUE, counts_weighted = FALSE,
counts_name = "counts", value_name = x, q1_name = "q1",
q3_name = "q3", ...)
</code></pre>

#### Arguments
data A data.frame.
x Variable name in data to summarize.
w Optional name of the column in data with case weights.
* ``by``: An optional vector of column names in data used to group the results.
stats Statistic to calculate: "mean" or "quartiles".
counts Should group counts be added?
counts_weighted
Should counts be weighted by the case weights? If TRUE, the sum of w is
returned by group.
* ``counts_name``:  Name of column in the resulting data.frame containing the counts.
value_name Name of the resulting column with mean or median.
q1_name Name of the resulting column with first quartile values. Only relevant for stats
"quartiles".
q3_name Name of the resulting column with third quartile values. Only relevant for stats
"quartiles".
... Additional arguments passed to MetricsWeighted::weighted_mean or weighted_quartiles.
is.flashlight 9

#### Value

A data.frame with columns by, x and optionally counts_name.

#### Examples
```{r}

grouped_stats(iris, "Sepal.Width")
grouped_stats(iris, "Sepal.Width", stats = "quartiles")
grouped_stats(iris, "Sepal.Width", w = "Petal.Width")
grouped_stats(iris, "Sepal.Width", w = "Petal.Width", counts_weighted = TRUE)
grouped_stats(iris, "Sepal.Width", by = "Species")
grouped_stats(iris, "Sepal.Width", stats = "quartiles", by = "Species")
grouped_stats(iris, "Sepal.Width", w = "Petal.Width", by = "Species")
grouped_stats(iris, "Sepal.Width", w = "Petal.Width",
counts_weighted = TRUE, by = "Species")
grouped_stats(iris, "Sepal.Width", counts = FALSE)
grouped_stats(iris, "Sepal.Width", counts_name = "n",
stats = "quartiles", q1_name = "p25", q3_name = "p75")
```
