---
title: "``flashlight::light_importance``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 Permutation Importance

#### Description

Calculates performance per variable with respect to a performance measure before and after permuting
its values. The difference is a measure of importance, see Fisher et al. 2018 [1].

#### Usage
<pre><code>

light_importance(x, ...)
## Default S3 method:
light_importance(x, ...)
## S3 method for class 'flashlight'
light_importance(x, data = x$data, by = x$by,
metric = x$metrics[1], v = NULL, n_max = Inf, seed = NULL,
lower_is_better = TRUE, use_linkinv = FALSE,
metric_name = "metric", value_name = "value", label_name = "label",
22 light_importance
variable_name = "variable", ...)
## S3 method for class 'multiflashlight'
light_importance(x, ...)
</code></pre>

#### Arguments
* ``x``: An object of class flashlight or multiflashlight.
... Further arguments passed to light_performance.
* ``data``: An optionaldata.frame.
* ``by``: An optional vector of column names used to additionally group the results.
* ``metric``: An optional namedlist of length one with a metric as element. Defaults to the first metric in the flashlight. The metric needs to be a function with at least four
arguments: actual, predicted, case weights w and ....
v Vector of variables to assess importance for. Defaults to all variables in data.
* ``n_max``:  Maximum number of rows to consider. Use if data is large.
seed An integer random seed used to select and shuffle rows.
lower_is_better
Logical flag indicating if lower values in the metric are better or not. If set to FALSE, the increase in metric is multiplied by -1.
use_linkinv Should retransformation function be applied? Default is FALSE.
metric_name Name of the resulting column containing the name of the metric. Defaults to`"metric".
value_name Column name in resulting data containing the drop in performance. Defaults to "value".
* ``label_name``: Column name in resulting data containing the label of the flashlight. Defaults to "label".
variable_name Column name in resulting data containing the variable names. Defaults to "variable".

#### Details

The minimum required elements in the (multi-) flashlight are "y", "``predict_function``", "model",
"data" and "metrics". The latter two can also directly be passed to light_importance. Note that
by default, no retransformation function is applied.

#### Value

An object of class light_importance, light (and a list) with the following elements.
*  data A tibble with results. Can be used to build fully customized visualizations. The columns
"value_original" and "value_shuffled" provide the performance before and after shuffling.
*  by Same as input by.
*  metric_name Column name representing the name of the metric. For information only.
*  value_name Same as input value_name.
*  label_name Same as input label_name.
*  variable_name Same as input variable_name.
light_performance 23
Methods (by class)
*  default: Default method not implemented yet.
*  flashlight: Variable importance for a flashlight.
*  multiflashlight: Variable importance for a multiflashlight.

#### References

[1] Fisher A., Rudin C., Dominici F. (2018). All Models are Wrong but many are Useful: Variable
Importance for Black-Box, Proprietary, or Misspecified Prediction Models, using Model Class
Reliance. ArXiv <arxiv.org/abs/1801.01489>.
See Also
most_important, plot.light_importance.

#### Examples
```{r}

fit_part <- lm(Sepal.Length ~ Petal.Length, data = iris)
fit_full <- lm(Sepal.Length ~ ., data = iris)
mod_full <- flashlight(model = fit_full, label = "full", data = iris, y = "Sepal.Length")
mod_part <- flashlight(model = fit_part, label = "part", data = iris, y = "Sepal.Length")
mods <- multiflashlight(list(mod_full, mod_part), by = "Species")
light_importance(mod_full)
light_importance(mods)
ir <- iris
ir$log_sl <- log(ir$Sepal.Length)
fit_lm <- lm(log_sl ~ Petal.Length, data = ir)
fit_glm <- glm(Sepal.Length ~ Petal.Length, data = ir, family = Gamma(link = log))
fl_lm <- flashlight(model = fit_lm, label = "lm", y = "log_sl", linkinv = exp)
fl_glm <- flashlight(model = fit_glm, label = "glm", y = "Sepal.Length",
``predict_function`` = function(m, X) predict(m, X, type = "response"))
fls <- multiflashlight(list(fl_lm, fl_glm), data = ir)
light_importance(fls, v = "Petal.Length", seed = 45)
light_importance(fls, v = "Petal.Length", seed = 45, use_linkinv = TRUE)
