---
title: "``flashlight::light_profile``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

light_profile Partial Dependence and other Profiles

#### Description

Calculates different types of profiles across covariable values. By default, partial dependence profiles
[1] are calculated. Other options are profiles of ALE (accumulated local effects, see [2]), response, predicted values ("M plots" or "marginal plots", see [2]) and residuals. The results are aggregated either by (weighted) means or by (weighted) quartiles. Note that ALE profiles are calibrated
by (weighted) average predictions. In contrast to the suggestions in [2], we calculate ALE profiles of factors in the same order as the factor levels. They are not being reordered based on
similiarity of other variables.

#### Usage
<pre><code>

light_profile(x, ...)
## Default S3 method:
light_profile(x, ...)
## S3 method for class 'flashlight'
light_profile(xNULL, data = NULL,
by = x$by, type = c("partial dependence", "ale", "predicted",
"response", "residual"), stats = c("mean", "quartiles"),
breaks = NULL, n_bins = 11, cut_type = c("equal", "quantile"),
26 light_profile
use_linkinv = TRUE, value_name = "value", q1_name = "q1",
q3_name = "q3", label_name = "label", type_name = "type",
counts_name = "counts", counts = TRUE, counts_weighted = FALSE,
v_labels = TRUE, pred = NULL, pd_evaluate_at = NULL,
pd_grid = NULL, pd_indices = NULL, pd_n_max = 1000,
pd_seed = NULL, pd_center = FALSE, ale_two_sided = FALSE, ...)
## S3 method for class 'multiflashlight'
light_profile(xNULL, data = NULL,
breaks = NULL, n_bins = 11, cut_type = c("equal", "quantile"),
pd_evaluate_at = NULL, pd_grid = NULL, ...)
</code></pre>

#### Arguments
* ``x``: An object of class flashlight or multiflashlight.
... Further arguments passed to cut3 resp. formatC in forming the cut breaks of
the v variable. Not relevant for partial dependence and ALE profiles.
v The variable to be profiled.
* ``data``: An optionaldata.frame.
* ``by``: An optional vector of column names used to additionally group the results.
type Type of the profile: Either "partial dependence", "ale", "predicted", "response",
or "residual".
stats Statistic to calculate: "mean" or "quartiles". For ALE profiles, only "mean"
makes sense.
breaks Cut breaks for a numeric v.
* ``n_bins``: Maxmium number of unique values to evaluate for numeric v. Only used if
neither grid nor pd_evaluate_at is specified.
cut_type For the default "equal", bins of equal width are created for v by pretty. Choose
"quantile" to create quantile bins.
use_linkinv Should retransformation function be applied? Default is TRUE.
value_name Column name in resulting data containing the profile value. Defaults to "value".
q1_name Name of the resulting column with first quartile values. Only relevant for stats
"quartiles".
q3_name Name of the resulting column with third quartile values. Only relevant for stats
"quartiles".
* ``label_name``: Column name in resulting data containing the label of the flashlight. Defaults
to "label".
type_name Column name in the resulting data with the plot type.
* ``counts_name``:  Name of the column containing counts if counts is TRUE.
counts Should counts be added?
counts_weighted
If counts is TRUE: Should counts be weighted by the case weights? If TRUE,
the sum of w is returned by group.
light_profile 27
v_labels If FALSE, return group centers of v instead of labels. Only relevant for types "response", "predicted" or "residual" and if v is being binned. In that case useful
if e.g. different flashlights use different data sets and bin labels would not match.
pred Optional vector with predictions (after application of inverse link). Can be used to avoid recalculation of predictions over and over if the functions is to be repeatedly
called for different v and predictions are computationally expensive to make. Only relevant for type = "predicted" and type = "ale".
pd_evaluate_at Vector with values of v used to evaluate the profile. Only relevant for type =
"partial dependence" and "ale".
pd_grid A data.frame with grid values, e.g. generated by expand.grid. Only used for type = "partial dependence".
pd_indices A vector of row numbers to consider in calculating partial dependence profiles.
Only used for type = "partial dependence" and "ale".
pd_n_max Maximum number of ICE profiles to calculate (will be randomly picked from
data). Only used for type = "partial dependence" and "ale".
pd_seed Integer random seed used to select ICE profiles. Only used for type = "partial dependence" and "ale".
pd_center Should ICE curves be centered within by subsets before caclulating partial dependence
profiles? This option is interesting together with stats = "quartiles"
in order to visualize interaction strength.
ale_two_sided If TRUE, v is continuous and breaks are passed or being calculated, then twosided
derivatives are calculated for ALE instead of left derivatives. More specifically:
Usually, local effects at value x are calculated using points between x-e
and x. Set ale_two_sided = TRUE to use points between x-e/2 and x+e/2.

#### Details

For numeric covariables v with more than n_bins disjoint values, its values are binned. Alternatively, breaks can be provided to specify the binning. For partial dependence profiles (and partly also ALE profiles), this behaviour can be overritten either by providing a vector of evaluation points (``pd_evaluate_at``) or an evaluation pd_grid. By the latter we mean a data frame with column name(s) with a (multi-)variate evaluation grid. For partial dependence, ALE and prediction profiles,
"model", "``predict_function``", linkinv" and "data" are required. For response profiles its just "y", "linkinv" and "data". "data" can be passed on the fly for both types.

#### Value

An object of classes light_profile, light (and a list) with the following elements.
*  data A tibble containing results. Can be used to build fully customized visualizations. Its
column names are specified by all other items in this list.
*  by Names of group by variable.
*  v The variable(s) evaluated.
*  type Same as input type. For information only.
*  stats Same as input stats.
28 light_profile
*  value_name Same as input value_name.
*  q1_name Same as input q1_name.
*  q3_name Same as input q3_name.
*  label_name Same as input label_name.
*  type_name Same as input type_name.
*  counts_name Same as input counts_name.
Methods (by class)
*  default: Default method not implemented yet.
*  flashlight: Profiles for flashlight.
*  multiflashlight: Profiles for multiflashlight.

#### References

[1] Friedman J. H. (2001). Greedy function approximation: A gradient boosting machine. The
Annals of Statistics, 29:1189–1232. [2] Apley D. W. (2016). Visualizing the effects of predictor
variables in black box supervised learning models. ArXiv <arXiv:1612.08468>.
See Also
light_effects, plot.light_profile.

#### Examples
```{r}

fit_full <- lm(Sepal.Length ~ ., data = iris)
fit_part <- lm(Sepal.Length ~ Petal.Length, data = iris)
mod_full <- flashlight(model = fit_full, label = "full", data = iris, y = "Sepal.Length")
mod_part <- flashlight(model = fit_part, label = "part", data = iris, y = "Sepal.Length")
mods <- multiflashlight(list(mod_full, mod_part))
light_profile(mod_full, v = "Species")
light_profile(mod_full, v = "Species", counts = FALSE)
light_profile(mod_full, v = "Species", type = "response")
light_profile(mod_full, v = "Species", type = "ale")
light_profile(mod_full, v = "Species", stats = "quartiles")
light_profile(mod_full, v = "Petal.Width")
light_profile(mod_full, v = "Petal.Width", type = "residual")
light_profile(mod_full, v = "Petal.Width", type = "residual", v_label = FALSE)
light_profile(mod_full, v = "Petal.Width", type = "residual", dig.lab = 1)
light_profile(mod_full, v = "Petal.Width", stats = "quartiles")
light_profile(mod_full, v = "Petal.Width", n_bins = 3)
light_profile(mod_full, v = "Petal.Width", pd_evaluate_at = 2:4)
light_profile(mod_full, pd_grid = data.frame(Petal.Width = 2:4))
light_profile(mod_full, v = "Petal.Width", by = "Species")
light_profile(mods, v = "Petal.Width")
light_recode 29
light_profile(mods, v = "Petal.Width", by = "Species")
light_profile(mods, v = "Petal.Width", by = "Species", type = "predicted")
light_profile(mods, v = "Petal.Width", by = "Species",
type = "predicted", stats = "quartiles")
light_profile(mods, v = "Petal.Width", by = "Species", stats = "quartiles",
value_name = "pd", q1_name = "p25", q3_name = "p75", label_name = "model",
type_name = "visualization", counts_name = "n")
```

