---
title: "``flashlight::light_effects``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 Combination of Response, Predicted, Partial Dependence, and ALE
Profiles

#### Description

Calculates response- prediction-, partial dependence, and ALE profiles of a (multi-)flashlight with
respect to a covariable v.

#### Usage
<pre><code>

light_effects(x, ...)
## Default S3 method:
light_effects(x, ...)
## S3 method for class 'flashlight'
light_effects(x, v, data = NULL, by = x$by,
stats = c("mean", "quartiles"), breaks = NULL, n_bins = 11,
cut_type = c("equal", "quantile"), use_linkinv = TRUE,
value_name = "value", q1_name = "q1", q3_name = "q3",
label_name = "label", type_name = "type", counts_name = "counts",
counts_weighted = FALSE, v_labels = TRUE, pred = NULL,
pd_indices = NULL, pd_n_max = 1000, pd_seed = NULL,
ale_two_sided = TRUE, ...)
## S3 method for class 'multiflashlight'
light_effects(x, v, data = NULL,
breaks = NULL, n_bins = 11, cut_type = c("equal", "quantile"), ...)
</code></pre>

#### Arguments
* ``x``: An object of class flashlight or multiflashlight.
... Further arguments passed to cut3 resp. formatC in forming the cut breaks of
the v variable.
v The variable to be profiled.
* ``data``: An optionaldata.frame.
* ``by``: An optional vector of column names used to additionally group the results.
stats Statistic to calculate for the response profile: "mean" or "quartiles".
breaks Cut breaks for a numeric v.
* ``n_bins``: Maxmium number of unique values to evaluate for numeric v.
cut_type For the default "equal", bins of equal width are created for v by pretty. Choose
"quantile" to create quantile bins (recommended if interested in ALE).
use_linkinv Should retransformation function be applied? Default is TRUE.
light_effects 17
value_name Column name in resulting data objects containing the profile value. Defaults to "value".
q1_name Name of the resulting column with first quartile values. Only relevant for stats "quartiles".
q3_name Name of the resulting column with third quartile values. Only relevant for stats "quartiles".
* ``label_name``: Column name in resulting data containing the label of the flashlight. Defaults to "label".
type_name Name of the column in data containing type.
* ``counts_name``:  Name of the column containing counts.
counts_weighted Should counts be weighted by the case weights? If TRUE, the sum of w is
returned by group.
v_labels If FALSE, return group centers of v instead of labels. Only relevant if v is
numeric with many distinct values. In that case useful if e.g. different flashlights
use different data sets.
pred Optional vector with predictions (after application of inverse link). Can be used
to avoid recalculation of predictions over and over if the functions is to be repeatedly
called for different v and predictions are computationally expensive to
make.
pd_indices A vector of row numbers to consider in calculating partial dependence and ALE
profiles. Useful to force all flashlights to use the same basis for calculations of
partial dependence and ALE.
pd_n_max Maximum number of ICE profiles to consider for partial depencence and ALE
calculation (will be randomly picked from data).
pd_seed An integer random seed used to sample ICE profiles for partial dependence and
ALE.
ale_two_sided If TRUE, v is continuous and breaks are passed or being calculated, then twosided
derivatives are calculated for ALE instead of left derivatives. This aligns
the results better with the x labels. More specifically: Usually, local effects at
value x are calculated using points between x-e and x. Set ale_two_sided =
TRUE to use points between x-e/2 and x+e/2.

#### Details

Note that ALE profiles are being calibrated by (weighted) average predictions. The resulting level
might be quite different from the one of the partial dependence profiles.

#### Value

An object of classes light_effects, light (and a list) with the following elements.
*  response A tibble containing the response profiles.
*  predicted A tibble containing the prediction profiles.
*  pd A tibble containing the partial dependence profiles.
18 light_effects
*  by Same as input by.
*  v The variable(s) evaluated.
*  stats Same as input stats.
*  value_name Same as input value_name.
*  q1_name Same as input q1_name.
*  q3_name Same as input q3_name.
*  label_name Same as input label_name.
*  type_name Same as input type.
*  counts_name Same as input counts_name.
Methods (by class)
*  default: Default method.
*  flashlight: Profiles for a flashlight object.
*  multiflashlight: Effect profiles for a multiflashlight object.
See Also
light_profile, plot.light_effects.

#### Examples
```{r}

fit_full <- lm(Sepal.Length ~ ., data = iris)
fit_part <- lm(Sepal.Length ~ Petal.Length, data = iris)
mod_full <- flashlight(model = fit_full, label = "full", data = iris, y = "Sepal.Length")
mod_part <- flashlight(model = fit_part, label = "part", data = iris, y = "Sepal.Length")
mods <- multiflashlight(list(mod_full, mod_part))
light_effects(mod_full, v = "Species")
light_effects(mod_full, v = "Species", stats = "quartiles")
light_effects(mod_full, v = "Petal.Width")
light_effects(mod_full, v = "Petal.Width", v_label = FALSE)
light_effects(mod_full, v = "Petal.Width", stats = "quartiles")
light_effects(mod_full, v = "Petal.Width", n_bins = 3)
light_effects(mod_full, v = "Petal.Width", n_bins = 3, format = "f")
light_effects(mod_full, v = "Petal.Width", breaks = 0:3)
light_effects(mod_full, v = "Petal.Width", by = "Species")
light_effects(mods, v = "Petal.Width")
light_effects(mods, v = "Petal.Width", by = "Species")
light_effects(mods, v = "Petal.Width", by = "Species", stats = "quartiles")
