---
title: "``flashlight::....``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

light_check Check flashlight

#### Description

Checks if an object of class flashlight or multiflashlight is consistently defined.
14 light_combine

#### Usage
<pre><code>

light_check(x)
## Default S3 method:
light_check(x)
## S3 method for class 'flashlight'
light_check(x)
## S3 method for class 'multiflashlight'
light_check(x)
</code></pre>

#### Arguments
* ``x``: An object of class flashlight or multiflashlight.

#### Value

The input x or an error message.
Methods (by class)
*  default: Default check method not implemented yet.
*  flashlight: Checks if a flashlight object is consistently defined.
*  multiflashlight: Checks if a multiflashlight object is consistently defined.

#### Examples
```{r}

fit <- lm(Sepal.Length ~ ., data = iris)
fit_log <- lm(log(Sepal.Length) ~ ., data = iris)
fl <- flashlight(fit, data = iris, y = "Sepal.Length", label = "ols")
fl_log <- flashlight(fit_log, y = "Sepal.Length", label = "ols", linkinv = exp)
light_check(fl)
light_check(fl_log)
```
