---
title: "``flashlight::summary.compare``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

multiflashlight Create or Update a multiflashlight

#### Description

Combines a list of flashlights to an object of class multiflashlight and/or updates a multiflashlight.

#### Usage
<pre><code>

multiflashlight(x, ...)
## Default S3 method:
multiflashlight(x, ...)
## S3 method for class 'flashlight'
multiflashlight(x, ...)
## S3 method for class 'list'
multiflashlight(x, ...)
## S3 method for class 'multiflashlight'
multiflashlight(x, ...)
</code></pre>

#### Arguments
* ``x``: An object of class multiflashlight, flashlight or a list of flashlights.
... Optional arguments in the flashlights to update, see examples.

#### Value

An object of class multiflashlight. This is a named list of flashlight objects.
Methods (by class)
*  default: Used to create a flashlight object. No x has to be passed in this case.
*  flashlight: Updates an existing flashlight object and turns into a multiflashlight.
*  list: Creates (and updates) a multiflashlight from a list of flashlights.
*  multiflashlight: Updates an object of class multiflashlight.
See Also
flashlight.
plot.light_breakdown 33

#### Examples
```{r}

fit_lm <- lm(Sepal.Length ~ ., data = iris)
fit_glm <- glm(Sepal.Length ~ ., family = Gamma(link = log), data = iris)
mod_lm <- flashlight(model = fit_lm, label = "lm")
mod_glm <- flashlight(model = fit_glm, label = "glm")
(mods <- multiflashlight(list(mod_lm, mod_glm)))
mods <- multiflashlight(list(mod_lm, mod_glm),
data = iris, by = "Species", y = "Sepal.Length")
mod_lm <- mods$lm
mod_lm
```

