---
title: "``flashlight::plot.light_ice``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plot.light_ice)
```

plot.light_ice Visualize ICE profiles

#### Description

Minimal visualization of an object of class light_ice as geom_line. The object returned is of
class ggplot and can be further customized.

#### Usage
<pre><code>

## S3 method for class 'light_ice'
plot(x, facet_scales = "fixed", rotate_x = FALSE,
...)
</code></pre>

#### Arguments
* ``x``: An object of class light_ice.
* ``facet_scales``: Scales argument passed to facet_wrap.
rotate_x Should x axis labels be rotated by 45 degrees? Default is FALSE.
... Further arguments passed to geom_line.
36 plot.light_ice

#### Details

Each observation is visualized by a line. The first "by" variable is represented by the color, a second "by" variable or a multiflashlight by facets.

#### Value

An object of class ggplot2.
Author(s)
Michael Mayer
See Also
light_ice.

#### Examples
```{r}

fit_full <- lm(Sepal.Length ~ ., data = iris)
fit_part <- lm(Sepal.Length ~ Petal.Length, data = iris)
mod_full <- flashlight(model = fit_full, label = "full", data = iris, y = "Sepal.Length")
mod_part <- flashlight(model = fit_part, label = "part", data = iris, y = "Sepal.Length")
mods <- multiflashlight(list(mod_full, mod_part))
grid <- expand.grid(Species = levels(iris$Species), Petal.Length = 2:4)
plot(light_ice(mod_full, v = "Species"), alpha = 0.2)
indices <- (1:15) * 10
plot(light_ice(mod_full, v = "Species"), rotate_x = TRUE)
plot(light_ice(mods, v = "Species", indices = indices))
plot(light_ice(mods, v = "Species", indices = indices, center = TRUE))
plot(light_ice(mods, v = "Petal.Width", n_bins = 5, indices = indices))
plot(light_ice(mods, v = "Petal.Width", by = "Species", n_bins = 5, indices = indices))
ir <- iris
ir$log_sl <- log(ir$Sepal.Length)
fit_lm <- lm(log_sl ~ Petal.Length + Petal.Width, data = ir)
fit_glm <- glm(Sepal.Length ~ Petal.Length + Petal.Width,
data = ir, family = Gamma(link = log))
fl_lm <- flashlight(model = fit_lm, label = "lm", y = "log_sl", linkinv = exp)
fl_glm <- flashlight(model = fit_glm, label = "glm", y = "Sepal.Length",
``predict_function`` = function(m, X) predict(m, X, type = "response"))
fls <- multiflashlight(list(fl_lm, fl_glm), data = ir)
plot(light_ice(fls, v = "Petal.Length", indices = indices))
plot(light_ice(fls, v = "Petal.Length", indices = indices, center = TRUE))
plot(light_ice(fls, v = "Petal.Length", indices = indices, by = "Species", center = TRUE))
plot(light_ice(fls, v = "Petal.Length", indices = indices, use_linkinv = FALSE))
```
