---
title: "``flashlight::plot_counts``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

plot_counts Add Counts to Effects Plot

#### Description

Add counts as bar plot on top of light_effects plot.

#### Usage
<pre><code>

plot_counts(p, x, text_size = 3, facet_scales = "free_x",
show_labels = TRUE, big.mark = "'", scientific = FALSE, ...)
plot_counts 41
</code></pre>

#### Arguments
p The result of plot.light_effects.
* ``x``: An object of class light_effects.
text_size Size of count labels.
* ``facet_scales``: Scales argument passed to facet_wrap.
show_labels Should counts be added as text?
big.mark Parameter passed to format. Default is "’".
scientific Parameter passed to format. Default is FALSE.
... Further arguments passed to geom_bar.

#### Details

Experimental. Uses package ggpubr to rearrange the figure. Thus, the resulting plot cannot be easily modified. Furthermore, adding counts only works if the legend in ``plot.light_effects`` is not placed on the left or right side of the plot. It has to be placed inside or at the bottom.

#### Value

An object of class ggplot2.
See Also
plot.light_effects.

#### Examples
```{r}

fit_full <- lm(Sepal.Length ~ ., data = iris)
fit_part <- glm(Sepal.Length ~ Petal.Length, data = iris)
mod_full <- flashlight(model = fit_full, label = "full", data = iris, y = "Sepal.Length")
mod_part <- flashlight(model = fit_part, label = "part", data = iris, y = "Sepal.Length")
mods <- multiflashlight(list(mod_full, mod_part))
x <- light_effects(mod_full, v = "Petal.Width", stats = "quartiles")
plot_counts(plot(x), x, width = 0.3, alpha = 0.2)
plot_counts(plot(x, zero_counts = FALSE), x, width = 0.3, alpha = 0.2)
plot_counts(plot(x), x, width = 0.3, alpha = 0.2, show_labels = FALSE)
plot_counts(plot(x, use = "response"), x, fill = "lightblue")
plot_counts(plot(x, use = "pd", show.legend = FALSE), x, fill = "lightblue")
```
