---
title: "``flashlight::flashlight``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

flashlight Create or Update a flashlight

#### Description

Creates or updates a flashlight object. If a flashlight is to be created, all arguments are optional except label. If a flashlight is to be updated, all arguments are optional up to x (the flashlight to be updated).

#### Usage

<pre><code>
flashlight(x, ...)
## Default S3 method:
flashlight(x, model = NULL, data = NULLNULL,
``predict_function`` = predict, linkinv = function(z) zNULL,
by = NULL, metrics = list(rmse = rmse), label = NULL, ...)
## S3 method for class 'flashlight'
flashlight(x, ...)
</code></pre>

#### Arguments

* ``x``: An object of class flashlight. If not provided, a new flashlight is created based on further input. Otherwise, x is updated based on further input.
... Arguments passed from or to other functions.
model A fitted model of any type. Most models require a customized ``predict_function``.
data A data.frame or tibble used as basis for calculations.
y Variable name of response.
``predict_function``
A real valued function with two arguments: A model and a data of the same
structure as data. Only the order of the two arguments matter, not their names.
linkinv An inverse transformation function applied after ``predict_function``.
w A variable name of case weights.
by A character vector with names of grouping variables.
metrics A named list of metrics. Here, a metric is a function with exactly four arguments:
actual, predicted, w (case weights) and ... like those in package
MetricsWeighted.
label Name of the flashlight. Required.

#### Value

An object of class flashlight (and list) containing each input (except x) as element.
8 grouped_stats
Methods (by class)
*  default: Used to create a flashlight object. No x has to be passed in this case.
*  flashlight: Used to update an existing flashlight object.
See Also
multiflashlight.

#### Examples
```{r}

fit <- lm(Sepal.Length ~ ., data = iris)
(fl <- flashlight(model = fit, data = iris, y = "Sepal.Length", label = "ols"))
(fl_updated <- flashlight(fl, linkinv = exp))


