---
title: "``flashlight::light_recode``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

light_recode Recode Factor Columns

#### Description

Recodes factor levels of columns in data slots of an object of class light.

#### Usage
<pre><code>

light_recode(x, ...)
## Default S3 method:
light_recode(x, ...)
## S3 method for class 'light'
light_recode(x, what, levels, labels, ...)
</code></pre>

#### Arguments
* ``x``: An object of class light.
... Further arguments passed to factor.
what Column identifier in x (not column name) to be recoded, e.g. "type_name",
"label_name".
levels Current levels/values of type_name column (in desired order).
labels New levels of type_name column in same order as levels.

#### Value

x with new factor levels of type_name column.
Methods (by class)
*  default: Default method not implemented yet.
*  light: Recoding factors in data slots of light object.
See Also
plot.light_effects.

#### Examples
```{r}

fit_full <- lm(Sepal.Length ~ ., data = iris)
fit_part <- lm(Sepal.Length ~ Petal.Length, data = iris)
mod_full <- flashlight(model = fit_full, label = "full", data = iris, y = "Sepal.Length")
mod_part <- flashlight(model = fit_part, label = "part", data = iris, y = "Sepal.Length")
mods <- multiflashlight(list(mod_full, mod_part))
eff <- light_effects(mods, v = "Species")
eff <- light_recode(eff, what = "type_name",
levels = c("response", "predicted", "partial dependence"),
labels = c("Observed", "Fitted", "Effect"))
plot(eff)
perf <- light_performance(mods)
perf <- light_recode(perf, what = "label_name",
levels = c("part", "full"), labels = c("simple", "complex"))
plot(perf)
```

