---
title: "``flashlight::light_breakdown ``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

light_breakdown 11
light_breakdown Variable Contribution Breakdown for Single Observation

#### Description

Calculates sequential additive variable contribution to the prediction of a single observation, see
Gosiewska and Biecek [1] and the details below.

#### Usage
<pre><code>

light_breakdown(x, ...)
## Default S3 method:
light_breakdown(x, ...)
## S3 method for class 'flashlight'
light_breakdown(x, new_obs, data = x$data,
by = x$byNULL, order_by_importance = TRUE, top_m = Inf,
n_max = Inf, seed = NULL, use_linkinv = FALSE,
after_name = "after", before_name = "before", label_name = "label",
variable_name = "variable", step_name = "step",
description_name = "description", digits = 3, ...)
## S3 method for class 'multiflashlight'
light_breakdown(x, ...)
</code></pre>

#### Arguments
* ``x``: An object of class flashlight or multiflashlight.
... Further arguments passed to prettyNum to format numbers in description text.
new_obs One single new observation to calculate variable attribution for. Needs to be a data.frame of same structure as data.
* ``data``: An optionaldata.frame.
* ``by``: An optional vector of column names used to filter data for rows with equal values in "by" variables as new_obs.
v Vector of variables to assess contribution for. Defaults to all except those specified
by "y", "w" and "by".
order_by_importance
Logical flag indicated if v should be ordered automatically based on individual contribution (see description).
* ``top_m``:  Maximum number of steps/variable contributions to be calculated.
* ``n_max``:  Maximum number of rows in data to consider. Set to lower value if data is
large.
seed An integer random seed used to shuffle rows if n_max is smaller than the number
of rows in data.
12 light_breakdown
use_linkinv Should retransformation function be applied? Default is FALSE.
after_name Column name in resulting data containing prediction after the step in step_name.
Defaults to "after".
before_name Column name in resulting data containing prediction before the step in step_name.
Defaults to "before".
* ``label_name``: Column name in resulting data containing the label of the flashlight. Defaults
to "label".
variable_name Column name in resulting data containing the variable names. Defaults to "variable".
step_name Column name in resulting data containing the step of the prediction process.
Defaults to "step".
description_name
Column name in resulting data containing the description text to be visualized.
Defaults to "description".
digits Passed to prettyNum to format numbers in description text.

#### Details

The breakdown algorithm works as follows: First, the visit order (x_1, ..., x_m) of the variables
v is specified. Then, in the query data, the column x_1 is set to the value of x_1 of the single
observation new_obs to be explained. The change in the (weighted) average prediction on data measures the contribution of x_1 on the prediction of new_obs. This procedure is iterated over all x_i until eventually, all rows in data are identical to new_obs. A complication with this approach
is that the visit order is relevant, at least for non-additive models. Ideally, the algorithm could be
repeated for all possible permutations of v and its results averaged per variable. This is basically what SHAP values do, see [1] for an explanation. Unfortunately, there is no efficient way to do this in a model agnostic way. Thus we use the short-cut described in [1]. The variables are sorted
by the size of their contribution in the same way as the breakdown algorithm but without iteration, i.e. starting from the original query data for each variable $x_i$. Note that the minimum required elements in the (multi-) flashlight are "y", "``predict_function``", "model", and "data". The latter can
also directly be passed to light_breakdown. Note that by default, no retransformation function is
applied.

#### Value

An object of class light_breakdown, light (and a list) with the following elements.
*  data A tibble with results. Can be used to build fully customized visualizations.
*  by Same as input by.
*  after_name Same as input after_name.
*  before_name Same as input before_name.
*  label_name Same as input label_name.
*  variable_name Same as input variable_name.
*  step_name Same as input step_name.
*  description_name Same as input description_name.
light_check 13
Methods (by class)
*  default: Default method not implemented yet.
*  flashlight: Variable attribution to single observation for a flashlight.
*  multiflashlight: Variable attribution to single observation for a multiflashlight.

#### References

[1] A. Gosiewska and P. Biecek (2019). IBREAKDOWN: Uncertainty of model explanations for
non-additive predictive models. ArXiv <arxiv.org/abs/1903.11420>.
See Also
plot.light_breakdown.

#### Examples
```{r}

fit_part <- lm(Sepal.Length ~ Petal.Length, data = iris)
fit_full <- lm(Sepal.Length ~ ., data = iris)
mod_full <- flashlight(model = fit_full, label = "full", data = iris, y = "Sepal.Length")
mod_part <- flashlight(model = fit_part, label = "part", data = iris, y = "Sepal.Length")
mods <- multiflashlight(list(mod_full, mod_part), by = "Species")
light_breakdown(mod_full, new_obs = iris[1, ])
light_breakdown(mods, new_obs = iris[1, ])
light_breakdown(mods, new_obs = iris[1, ], top_m = 2)
ir <- iris
ir$log_sl <- log(ir$Sepal.Length)
fit_lm <- lm(log_sl ~ Petal.Length, data = ir)
fit_glm <- glm(Sepal.Length ~ Petal.Length, data = ir, family = Gamma(link = log))
fl_lm <- flashlight(model = fit_lm, label = "lm", y = "log_sl", linkinv = exp)
fl_glm <- flashlight(model = fit_glm, label = "glm", y = "Sepal.Length",
``predict_function`` = function(m, X) predict(m, X, type = "response"))
fls <- multiflashlight(list(fl_lm, fl_glm), data = ir)
light_breakdown(fls, new_obs = ir[1, ])$data
light_breakdown(fls, new_obs = ir[1, ], use_linkinv = TRUE)$data
```

