---
title: "``flashlight::light_performance``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


light_performance Model Performance of Flashlight

#### Description

Calculates performance of a flashlight with respect to one or more performance measure.
24 light_performance

#### Usage
<pre><code>

light_performance(x, ...)
## Default S3 method:
light_performance(x, ...)
## S3 method for class 'flashlight'
light_performance(x, data = x$data, by = x$by,
metrics = x$metrics, use_linkinv = FALSE, metric_name = "metric",
value_name = "value", label_name = "label", ...)
## S3 method for class 'multiflashlight'
light_performance(x, ...)
</code></pre>

#### Arguments
* ``x``: An object of class flashlight or multiflashlight.
... Arguments passed from or to other functions.
* ``data``: An optionaldata.frame.
* ``by``: An optional vector of column names used to additionally group the results. Will
overwrite x$by.
metrics An optional named list with metrics. Each metric takes at least four arguments:
actual, predicted, case weights w and ....
use_linkinv Should retransformation function be applied? Default is FALSE.
metric_name Column name in resulting data containing the name of the metric. Defaults to "metric".
value_name Column name in resulting data containing the value of the metric. Defaults to "value".
* ``label_name``: Column name in resulting data containing the label of the flashlight. Defaults to "label".

#### Details

The minimal required elements in the (multi-) flashlight are "y", "``predict_function``", "model", "data"
and "metrics". The latter two can also directly be passed to light_performance. Note that by
default, no retransformation function is applied.

#### Value

An object of class light_performance, light (and a list) with the following elements.
*  data A tibble containing the results. Can be used to build fully customized visualizations.
*  by Same as input by.
*  metric_name Same as input metric_name.
*  value_name Same as input value_name.
*  label_name Same as input label_name.

Methods (by class)
*  default: Default method not implemented yet.
*  flashlight: Model performance of flashlight object.
*  multiflashlight: Model performance of multiflashlight object.
See Also
plot.light_performance.

#### Examples
```{r}

fit_full <- lm(Sepal.Length ~ ., data = iris)
fit_part <- lm(Sepal.Length ~ Petal.Length, data = iris)
mod_full <- flashlight(model = fit_full, label = "full", data = iris, y = "Sepal.Length")
mod_part <- flashlight(model = fit_part, label = "part", data = iris, y = "Sepal.Length")
mods <- multiflashlight(list(mod_full, mod_part), by = "Species")
light_performance(mod_full)
light_performance(mod_full, metric_name = "perf",
value_name = "rmse", label_name = "model")
light_performance(mod_full, by = "Species")
light_performance(mods, by = "Species")
```
