---
title: "``flashlight::....``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

light_combine Combine Objects

#### Description

Combines a list of similar objects each of class light by row binding data.frame slots and retaining the other slots from the first list element.

#### Usage
<pre><code>

light_combine(x, ...)
## Default S3 method:
light_combine(x, ...)
## S3 method for class 'light'
light_combine(x, new_class = NULL, ...)
## S3 method for class 'list'
light_combine(x, new_class = NULL, ...)
</code></pre>

#### Arguments
x A list of objects of the same class.
... Further arguments passed from or to other methods.
new_class An optional vector with additional class names to be added to the output.

#### Value

If x is a list, an object like each element but with unioned rows in data slots.
Methods (by class)
*  default: Default method not implemented yet.
*  light: Since there is nothing to combine, the input is returned except for additional classes.
*  list: Combine a list of similar light objects.

#### Examples
```{r}

fit_lm <- lm(Sepal.Length ~ ., data = iris)
fit_glm <- glm(Sepal.Length ~ ., family = Gamma(link = "log"), data = iris)
mod_lm <- flashlight(model = fit_lm, label = "lm", data = iris, y = "Sepal.Length")
mod_glm <- flashlight(model = fit_glm, label = "glm", data = iris, y = "Sepal.Length",
``predict_function`` = function(object, newdata)
predict(object, newdata, type = "response"))
mods <- multiflashlight(list(mod_lm, mod_glm))
perf_lm <- light_performance(mod_lm)
perf_glm <- light_performance(mod_glm)
manual_comb <- light_combine(list(perf_lm, perf_glm),
new_class = "light_performance_multi")
auto_comb <- light_performance(mods)
all.equal(manual_comb, auto_comb)
light_combine(perf_lm)
```
