---
title: "``flashlight::summary.compare``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

most_important Most Important Variables.

#### Description

Returns the most important variable names sorted descendingly.

#### Usage
<pre><code>

most_important(x, top_m = Inf)
## Default S3 method:
most_important(x, top_m = Inf)
## S3 method for class 'light_importance'
most_important(x, top_m = Inf)
</code></pre>

#### Arguments
* ``x``: An object of class light_importance.
* ``top_m``:  Maximum number of important variables to be returned. Defaults to Inf, i.e.
return all variables in descending order of importance.

#### Value

A character vector of variable names sorted in descending order by importance.
Methods (by class)
*  default: Default method not implemented yet.
*  light_importance: Extracts most important variables from an object of class light_importance.
See Also
light_importance.

#### Examples
```{r}

fit <- lm(Sepal.Length ~ ., data = iris)
fl <- flashlight(model = fit, label = "ols", data = iris, y = "Sepal.Length")
(imp <- light_importance(fl, seed = 4))
most_important(imp)
most_important(imp, 2)
```
