---
title: "``flashlight::plot.light_profile``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(flashlight)
```

Visualize Profiles, e.g. of Partial Dependence

#### Description

Minimal visualization of an object of class light_profile. The object returned is of class ggplot and can be further customized.

#### Usage
<pre><code>

## S3 method for class 'light_profile'
plot(x, swap_dim = FALSE,
facet_scales = "free_x", rotate_x = x$type != "partial dependence",
...)
</code></pre>

#### Arguments
* ``x``: An object of class light_profile.
swap_dim If multiflashlight and one "by" variable or single flashlight with two "by" variables,
swap the role of dodge/fill variable and facet variable. If multiflashlight
or one "by" variable, use facets instead of colors.
* ``facet_scales``: Scales argument passed to facet_wrap.
rotate_x Should x axis labels be rotated by 45 degrees? TRUE, except for type "partial
dependence".
... Further arguments passed to geom_point and geom_line.



#### Examples
```{r}

fit_full <- lm(Sepal.Length ~ ., data = iris)
fit_part <- lm(Sepal.Length ~ Petal.Length, data = iris)
mod_full <- flashlight(model = fit_full, label = "full", data = iris, y = "Sepal.Length")
mod_part <- flashlight(model = fit_part, label = "part", data = iris, y = "Sepal.Length")
mods <- multiflashlight(list(mod_full, mod_part))
plot(light_profile(mod_full, v = "Species"))
plot(light_profile(mod_full, v = "Species", type = "ale"))
plot(light_profile(mod_full, v = "Species", type = "response"))
plot(light_profile(mod_full, v = "Species", type = "residual", stats = "quartiles"))
plot(light_profile(mod_full, v = "Petal.Width"))
plot(light_profile(mod_full, v = "Petal.Width", type = "residual"))
plot(light_profile(mod_full, v = "Petal.Width", stats = "quartiles"))
plot(light_profile(mod_full, v = "Petal.Width", by = "Species"))
plot(light_profile(mod_full, v = "Petal.Width", by = "Species"), swap_dim = TRUE)
plot(light_profile(mods, v = "Species"))
plot(light_profile(mods, v = "Petal.Width"))
plot(light_profile(mods, v = "Petal.Width"), swap_dim = TRUE)
plot(light_profile(mods, v = "Petal.Width", by = "Species"))
plot(light_profile(mods, v = "Petal.Width", by = "Species", type = "ale"))
plot(light_profile(mods, v = "Petal.Width", by = "Species"), swap_dim = TRUE)
plot(light_profile(mods, v = "Petal.Width", by = "Species", type = "predicted"))
plot(light_profile(mods, v = "Petal.Width", by = "Species",
type = "predicted", n_bins = 5), swap_dim = TRUE)
plot(light_profile(mods, v = "Petal.Width", by = "Species",
type = "predicted", stats = "quartiles"), rotate_x = TRUE)
```
#### Details

Either lines and points are plotted (if stats = "mean") or quartile boxes. If there is a "by" variable
or a multiflashlight, this first dimension is taken care by color (or if swap_dim = TRUE by facets).
If there are two "by" variables or a multiflashlight with one "by" variable, the first "by" variable is visualized as color, the second one or the multiflashlight via facet (change with swap_dim).
40 plot_counts

#### Value

An object of class ggplot2.
See Also
light_profile, plot.light_effects.
