---
title: "``flashlight::plot.light_importance ``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

37
plot.light_importance Visualize Model Importance

#### Description

Minimal visualization of an object of class light_importance as geom_bar. The object returned
is of class ggplot and can be further customized.

#### Usage
<pre><code>

## S3 method for class 'light_importance'
plot(x, top_m = Inf, swap_dim = FALSE,
facet_scales = "fixed", ...)
</code></pre>

#### Arguments
* ``x``: An object of class light_importance.
* ``top_m``:  Maximum number of important variables to be returned.
swap_dim If multiflashlight and one "by" variable or single flashlight with two "by" variables,
swap the role of dodge/fill variable and facet variable. If multiflashlight
or one "by" variable, use facets instead of colors.
* ``facet_scales``: Scales argument passed to facet_wrap.
... Further arguments passed to geom_bar.

#### Details

The plot is organized as a bar plot with variable names as x-aesthetic. Up to two additional dimensions (multiflashlight and one "by" variable or single flashlight with two "by" variables) can be visualized by facetting and dodge/fill. Set swap_dim = FALSE to revert the role of these two dimensions.
One single additional dimension is visualized by a facet wrap, or - if swap_dim = FALSE - by
dodge/fill.

#### Value

An object of class ggplot2.
See Also
light_importance.

#### Examples
```{r}

fit_full <- lm(Sepal.Length ~ ., data = iris)
fit_part <- lm(Sepal.Length ~ Petal.Length, data = iris)
mod_full <- flashlight(model = fit_full, label = "full", data = iris, y = "Sepal.Length")
mod_part <- flashlight(model = fit_part, label = "part", data = iris, y = "Sepal.Length")
mods <- multiflashlight(list(mod_full, mod_part), by = "Species")
38 plot.light_performance
plot(light_importance(mod_full), fill = "darkred")
plot(light_importance(mod_full, variable_name = "v", label_name = "model",
metric_name = "m", value_name = "drop"))
plot(light_importance(mod_full), top_m = 2)
plot(light_importance(mods))
plot(light_importance(mods), swap_dim = TRUE)
plot(light_importance(mods, by = NULL), fill = "darkgreen")
plot(light_importance(mods, by = NULL), swap_dim = TRUE)
plot.light_performance
```
