---
title: "``flashlight::Visualize``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Visualize Model Performance

#### Description

Minimal visualization of an object of class light_performance as geom_bar. The object returned
has class ggplot and can be further customized.

#### Usage
<pre><code>

## S3 method for class 'light_performance'
plot(x, swap_dim = FALSE,
facet_scales = "free_y", rotate_x = FALSE, ...)
</code></pre>

#### Arguments
* ``x``: An object of class light_performance.
swap_dim Should representation of dimensions (either two "by" variables or one "by" variable
and multiflashlight) of x aesthetic and dodge fill aesthetic be swapped?
Default is FALSE.
* ``facet_scales``: Scales argument passed to facet_wrap.
rotate_x Should x axis labels be rotated by 45 degrees? Default is FALSE.
... Further arguments passed to geom_bar.

#### Details

The plot is organized as a bar plot as follows: For flashlights without "by" variable specified, a single bar is drawn. Otherwise, the "by" variable (or the flashlight label if there is no "by" variable) is represented by the "x" aesthetic. The flashlight label (in case of one "by" variable) is represented
by dodged bars. This strategy makes sure that performance of different flashlights can be compared easiest. Set "swap_dim = TRUE" to revert the role of dodging and x aesthetic. Different metrics are
always represented by facets.

#### Value

An object of class ggplot2.
plot.light_profile 39
See Also
light_performance.

#### Examples
```{r}

fit_full <- lm(Sepal.Length ~ ., data = iris)
fit_part <- lm(Sepal.Length ~ Petal.Length, data = iris)
mod_full <- flashlight(model = fit_full, label = "full", data = iris, y = "Sepal.Length")
mod_part <- flashlight(model = fit_part, label = "part", data = iris, y = "Sepal.Length")
mods <- multiflashlight(list(mod_full, mod_part))
plot(light_performance(mod_full), fill = "darkred")
plot(light_performance(mod_full, by = "Species"), fill = "darkred")
plot(light_performance(mods))
plot(light_performance(mods, by = "Species"))
plot(light_performance(mods, by = "Species"), swap_dim = TRUE)
```
