documentclass[a4]{beamer}
usepackage{amssymb}
usepackage{graphicx}
usepackage{subfigure}
usepackage{framed}
usepackage{newlfont}
usepackage{amsmath,amsthm,amsfonts}
%usepackage{beamerthemesplit}
usepackage{pgf,pgfarrows,pgfnodes,pgfautomata,pgfheaps,pgfshade}
usepackage{mathptmx}  % Font Family
usepackage{helvet}   % Font Family
usepackage{color}

mode<presentation> {
 usetheme{Frankfurt} % was
 useinnertheme{rounded}
 useoutertheme{infolines}
 usefonttheme{serif}
 %usecolortheme{wolverine}
% usecolortheme{rose}
usefonttheme{structurebold}
}

setbeamercovered{dynamic}

title[MA4603]{Science Maths 3  {normalsize MA4603 Lecture 11A}}
author[Kevin O'Brien]{Kevin O'Brien  {scriptsize Kevin.obrien@ul.ie}}
date{Autumn Semester 2017}
institute[Maths & Stats]{Dept. of Mathematics & Statistics,  University textit{of} Limerick}

renewcommand{arraystretch}{1.5}

begin{document}





begin{frame}
frametitle{Shapiro-Wilk Test(a)}


begin{ }
 *           We will often be required to determine whether or not a data set is normally distributed.
 *           This assumption underpins many statistical models.
 *           The null hypothesis is that the data set is normally distributed.
 *           The alternative hypothesis is that the data set is not normally distributed.
 *           One procedure for testing these hypotheses is the Shapiro-Wilk test, implemented in texttt{R} using the command texttt{shapiro.test()}.
 *           (Remark: You will not be required to compute the test statistic for this test.)
end{ }
end{frame}
%----------------------------------------%
begin{frame}[fragile]
	frametitle{Shapiro-Wilk Test(b)}

textbf{IMPORTANT}
begin{ }
	 *           $H_0$ : The sample is drawn from a population that is normally distributed. 
	 *           $H_1$ : The sample is drawn from a population that is NOT normally distributed. 
end{ }
	
end{frame}
%----------------------------------------%
begin{frame}[fragile]
frametitle{Shapiro-Wilk Test(c)}
For the data set used previously; $x$ and $y$, we use the Shapiro-Wilk test to determine that both data sets are normally distributed.
<code>

> shapiro.test(X)

        Shapiro-Wilk normality test

data:  x
W = 0.9474, p-value = 0.6378

> shapiro.test(y)

        Shapiro-Wilk normality test

data:  y
W = 0.9347, p-value = 0.5273
</code>
end{frame}
%----------------------------------------%
begin{frame}[fragile]
	frametitle{Shapiro-Wilk Test(c)}
	
begin{ }
	 *           If we fail to reject $H_0$ we are saying that there is not enough evidence to contradict the null hypothesis. smallskip
	 *           This is not the same as saying that the data is a sample drawn from a normally distributioned population. smallskip
	 *           This is a common misconception.
	 *           Hypothesis Testing is about strength of evidence, not proof.
end{ }
	
	
end{frame}


%-------------------------------------------------%
begin{frame}[fragile]
frametitle{Graphical Procedures for Assessing Normality}

begin{ }
 *           The normal probability (Q-Q) plot is a very useful tool for determining whether or not a data set is normally distributed.
 *           textbf{Important} Interpretation is simple. If the points follow the trendline, we can assume that that data set is a sample drawn from a normally distributed population.
%(provided by the second line of texttt{R} code texttt{qqline}).
 *           One should expect minor deviations. Numerous major deviations would lead the analyst to conclude that the data set is not normally distributed.
 *           The Q-Q plot is best used in conjunction with a formal procedure such as the Shapiro-Wilk test.
end{ }

<pre>
<code>
>qqnorm(mySample)
>qqline(mySample)
</code>
</pre>
<p>
end{frame}

%-------------------------------------------------%

begin{frame}
frametitle{Graphical Procedures for Assessing Normality}

begin{center}
includegraphics[scale=0.32]{10AQQplot}
end{center}
end{frame}


begin{frame}
	frametitle{Interpretation: Can NOT Assume Normal Distribution}
	
	begin{center}
		includegraphics[scale=0.6]{QQplot2}
	end{center}
end{frame}


begin{frame}
	frametitle{Graphical Procedures for Assessing Normality}
	
	begin{center}
		includegraphics[scale=0.5]{QQplot1}
	end{center}
begin{ }
	 *           Can assume textbf{Verbal SAT} is a normally distributed variable.
	 *           Can not assume textbf{GPA} is a normally distributed variable.
end{ }
end{frame}

end{document}
